Nonlinear Tidal Dissipation in Binary White Dwarfs

### 2019-03-06 proposal/FINESST/Su_FINESST_2019.tex
LISA abbrev
clean up objectives

Compact white dwarf (WD) binary systems, with orbital periods in the range of
minutes to hours, are important for a range of astrophysical problems. They are
the most important sources of gravitational waves (GWs) for the Laser
Interferometric Space Antenna (LISA)\cite{lisa}. They are also thought to
produce interesting optical transients such as underluminous
supernovae\cite{underlum}, Ca-rich fast transients\cite{carich}, and tidal
novae\cite{tidal_novae}. Most importantly, they have been proposed as the likely
progenitors of type Ia supernovae (e.g.~\cite{Ia0,webbink} or more
recently\cite{Ia1,Ia2}). While presently only a few tens of compact WD binaries
are known\cite{lsst_wd}, \emph{Gaia} (currently gathering data) is expected to
expand the catalog to a few hundreds\cite{lsst_wd} (results based on
\emph{Gaia}'s second data release have already begun to
appear\cite{gaiaDD,gaiaDD2}), and the Large Synoptic Survey Telescope (LSST,
first light scheduled for 2020) will likely detect a few thousand
more\cite{lsst_wd}. These observations will significantly advance the
understanding of WD binaries and their evolution. My proposed theoretical and
computational research is well-timed to take advantage of these new advances.

@@

This project is extremely relevant to the NASA Astrophysics research program.
My work directly relates to (i) the interactions of particles under the extreme
conditions found in astrophysical situations, (ii) how complex systems create
and shape the structure and composition of the universe on all scales, and (iii)
the development of new techniques that can be applied to future major missions.

My work will improve understanding of possible energetic phenomena that can
occur in compact object binaries. My results concerning angular momentum
transfer via IGW breaking will be applicable to astrophysical systems beyond WD
binaries, for instance angular momentum transfer in massive
stars\cite{l_trans_rev}. I will devise new ways of analyzing astrophysically
interesting systems in \emph{Gaia} and LSST\@. Finally, my GW templates will be
important for LISA GW detection efforts.

My work has direct relevance to NASA missions in the detectability of
astrophysical transients related to WD binaries. For instance, tidal novae are
theorized to have a similar observational signature to dwarf novae which have
been observed with Chandra and the Hubble Space Telescope, among others.

@@

\end{document}
% mesa school?
% finish FINESST
% e disp? plot

% Kushnir?
% Bouss?? maybe

>>>>

Compact white dwarf (WD) binary systems, with orbital periods in the range of
minutes to hours, are important for a range of astrophysical problems. They are
the most important sources of gravitational waves (GWs) for the Laser
Interferometer Space Antenna (LISA)\cite{lisa}. They are also thought to
produce interesting optical transients such as underluminous
supernovae\cite{underlum}, Ca-rich fast transients\cite{carich}, and tidal
novae\cite{tidal_novae}. Most importantly, they have been proposed as the likely
progenitors of type Ia supernovae (e.g.~\cite{Ia0,webbink} or more
recently\cite{Ia1,Ia2}). While presently only a few tens of compact WD binaries
are known\cite{lsst_wd}, \emph{Gaia} (currently gathering data) is expected to
expand the catalog to a few hundreds\cite{lsst_wd} (results based on
\emph{Gaia}'s second data release have already begun to
appear\cite{gaiaDD,gaiaDD2}), and the Large Synoptic Survey Telescope (LSST,
first light scheduled for 2020) will likely detect a few thousand
more\cite{lsst_wd}. These observations will significantly advance the
understanding of WD binaries and their evolution. My proposed theoretical and
computational research is well-timed to take advantage of these new advances.

@@

This project is extremely relevant to the NASA Astrophysics Division, both in
terms of supporting the division's objectives and enhancing the science output
from NASA missions. My work directly relates to (i) the interactions of
particles under the extreme conditions found in astrophysical situations, (ii)
how complex systems create and shape the structure and composition of the
universe on all scales, and (iii) the development of new techniques that can be
applied to future major missions.

Towards (i), my work will improve understanding of possible energetic phenomena
that can occur in compact object binaries. Towards (ii), my results concerning
angular momentum transfer via IGW breaking will be applicable to astrophysical
systems beyond WD binaries, for instance in massive stars\cite{l_trans_rev}.
Towards (iii), my work has direct relevance to NASA missions by forecasting
astrophysical transients related to WD binaries. For instance, tidal novae are
theorized to have a similar observational signature to dwarf novae which have
been observed with Chandra and the Hubble Space Telescope, among others.
Additionally, my GW templates will also be important for LISA GW detection
efforts.

@@

\end{document}
